\documentclass[12pt, twoside]{report}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage[a4paper,width=150mm,top=25mm,bottom=25mm,bindingoffset=6mm]{geometry}
%\usepackage{biblatex}
\usepackage{setspace}
\usepackage[english]{babel} 
%\usepackage{natbib}
\usepackage{csquotes}
\usepackage[style=authoryear, backend=biber]{biblatex}
%\DeclareLanguageMapping{american}{american-apa}
%\usepackage{apacite}
\addbibresource{references.bib}
%\bibliographystyle{plainnat}
%\setcitestyle{authoryear,open={(},close={)}}

\graphicspath{ {images/} }

\linespread{1.25}


\begin{document}
\input{title.tex}

\chapter*{Abstract}
Understanding gradable adjectives like “big” requires making reference to a so-called comparison class - a set of objects the referent is implicitly compared to. For example, the utterance ‘“That Great Dane is big” could mean “That Great Dane is big compared to dogs in general” or “That Great Dane is big compared to other Great Danes”; yet the comparison class is rarely stated explicitly. So how do listeners establish the comparison class, given multiple a priori reasonable options?
Research on gradable adjectives has focused on the representation and integration of comparison classes into compositional semantics, but little is known about how human listeners decide upon a comparison class. 
This work takes a functional perspective on comparison class inference, guided by informational goals that speakers pursue when producing an utterance with a gradable adjective, and how listeners expect these goals to be achieved syntactically. For instance, given simple “Subject Predicate” sentences listeners expect that the subject aids reference (i.e., identifies the target), whereas the predicate accomplishes predication (i.e. asserts a property of the subject). Therefore, the role of the noun for comparison class determination  in a simple “Subject Predicate” sentence depends on its syntactic position. A noun appearing in the predicate is more likely to be intended to constrain the comparison class, whereas a noun in the subject can be explained away as intended for reference, leaving comparison class inference to other pragmatic reasoning. 
Converging evidence from four behavioural experiments supporting this proposal is presented alongside a novel formalisation of the inferential account in a qualitative computational model within the Rational Speech Act framework. This work contributes to the body of research on gradable adjectives, and provides a case study of  context-dependent language, emphasizing the complexity of the relation between the form and the meaning of linguistic expressions. 


\chapter*{Acknowledgements}
I want to thank...

\tableofcontents
\listoffigures
\listoftables

\chapter{Introduction}
\input{chapter01}

\chapter{Understanding Gradable Adjectives}
\input{chapter02}

\chapter{A Fucntional Perspective on Comprison Class Inference}
\input{chapter03}

\chapter{Experiments}
\input{chapter04}

\chapter{A Bayesian Reference-Predication Model}
\input{chapter05}

\chapter{Discussion}
\input{chapter06}

\chapter*{Declaration}
I declare that..

\appendix
\chapter{Appendix}	
\input{appendix}

\printbibliography
%\bibliography{references}
\end{document}
